---
title: "Connect R to Remote MySQL Instance"
output: html_notebook
---

## MySQL on AWS

This tutorial explains how to connect to a RDS on AWS. The RDS is a (free) MySQL database instance that must be set up prior through [AWS](https://aws.amazon.com/rds/free/).

Requires the **RMySQL** package to be installed.

```{r}
# 1. Library
library(RMySQL)

# 2. Settings
db_user <- 'user'
db_password <- 'password'
db_name <- 'cs5200-db-1'
db_table <- 'courses'
db_host <- 'cs5200-db-1.c07w5dbka9x.us-east-2.rds.amazonaws.com' # AWS Host
db_port <- 3306

# 3. Read data from db
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
s <- paste0("select * from ", db_table)
rs <- dbSendQuery(mydb, s)
df <-  fetch(rs, n = -1)
df
```

## MySQL on db4free.net

```{r}
# 1. Library
library(RMySQL)

# 2. Settings
db_user <- 'cs5200' 
db_password <- 'khour#2021'
db_name <- 'cs5200db1'
db_table <- 'courses'
db_host <- 'db4free.net' # Host 85.10.205.173
db_port <- 3306

# 3. Read data from db
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

```{sql connection=mydb}
create table courses (
  cid INTEGER NOT NULL,
  title varchar(32) NOT NULL,
  credits INTEGER NOT NULL
)
```

```{sql connection=mydb}
insert into courses
  values (1100,"CS and Application", 4)
```

```{sql connection=mydb}
select * from courses
```

```{r}
s <- paste0("select * from ", "courses")
rs <- dbSendQuery(mydb, s)
df <-  fetch(rs, n = -1)
df
```


## Disconnect from Database

Disconnect from the database (important as database servers have limited numbers of connection and each connection uses a resource).

```{r}
dbDisconnect(mydb)
```

